---
title: "Project 2 Stat4360"
output: pdf_document
date: "2024-02-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
wine <- read.table("~/wine.txt", header = T, sep = '')
#View(wine)
```

```{r}
head(wine)
attach(wine)
```
Question 1:
(a)  Perform an exploratory analysis of data. Comment on findings that interest you.
```{r}
# make the full model of the data
full <- lm(Quality~Clarity + Aroma + Body + Flavor + Oakiness + Region)
summary(full)
# make a reduced model from the full model
reduce <- lm(Quality~Aroma + Flavor + Oakiness)
summary(reduce)
# confidence interval
confint(reduce, level = 0.95)
data <- data.frame(Aroma = 3.4, Flavor = 3.2, Oakiness = 4.2)
predict(reduce, newData = data, interval = 'predict') 
# make a correlation
correlation = cor(wine[,c("Quality", "Clarity", "Aroma", "Body", "Flavor", "Oakiness")])
print(correlation)
# make a scatter plot
pairs(wine[, -7], main = " Scatterplots")
summary(wine)
```
(b) Do part (a) of Exercise 15 in Chapter 3 for these data.
For each predictor, fit a simple linear regression model to predict
the response. Describe your results. In which of the models is
there a statistically significant association between the predictor
and the response? Create some plots to back up your assertions.
```{r}
predictors <- c("Clarity", "Aroma", "Body", "Flavor", "Oakiness", "Region")
significant_predictors <- c()
# making the plots to find a correlation
for (predictor in predictors) {
  model <- lm(Quality ~ get(predictor), data = wine)
  summary_model <- summary(model)
  
  # Display results
  cat("\n\n=== Simple Linear Regression for", predictor, "===\n")
  print(summary_model)
  
  # Check for statistical significance (p-value < 0.05)
  if (summary_model$coefficients[2, "Pr(>|t|)"] < 0.05) {
    significant_predictors <- c(significant_predictors, predictor)
  }
  
  # Create scatter plots
  plot(wine[[predictor]], wine$Quality, main = paste("Scatter plot for", predictor),
       xlab = predictor, ylab = "Quality")
  abline(model, col = "red")
}
Quality_Aroma = lm(wine$Quality ~ wine$Aroma)
summary(Quality_Aroma)
Quality_Body = lm(wine$Quality ~ wine$Body)
summary(Quality_Body)
Quality_Flavor = lm(wine$Quality ~ wine$Flavor)
summary(Quality_Flavor)
Quality_Oakiness = lm(wine$Quality ~ wine$Oakiness)
summary(Quality_Oakiness)
Quality_Region = lm(wine$Quality ~ wine$Region)
summary(Quality_Region)

```
(c) Do part (b) of Exercise 15 in Chapter 3 for these data.
Fit a multiple regression model to predict the response using
all of the predictors. Describe your results. For which predictors
can we reject the null hypothesis H0 : βj = 0?
```{r}
# Multiple Regression Model
multiple_model <- lm(Quality ~ Clarity + Aroma + Body + Flavor + Oakiness + Region, data = wine)
summary_multiple_model <- summary(multiple_model)

# Display results for multiple regression
cat("\nMultiple Regression Model\n")
print(summary_multiple_model)

# Check for significant predictors (p-value < 0.05)
significant_predictors_multiple <- names(which(summary_multiple_model$coefficients[, "Pr(>|t|)"] < 0.05))

# Diagnostic plots for multiple regression
par(mfrow = c(2, 2))
plot(multiple_model)

# Display significant predictors in the multiple regression model
cat("\nSignificant predictors in the multiple regression model:", paste(significant_predictors_multiple, collapse = ", "), "\n")
# because less than 0.05 i got flavor and oakiness as the two significant predictions both having a positive direction of association and I can reject the null hypothesis for these two predictors
```
(d) Based on your observation in (b) and (c), build a “reasonably good” multiple
regression model for these data. Be sure to explore interactions of Region with 
other predictors. Carefully justify all the choices you make in building the 
model and verify the model assumptions.
```{r}
reduced_model <- lm(Quality~Aroma + Flavor + Oakiness)
summary(reduced_model)
confint(reduced_model, level = 0.95)
df = data.frame(Aroma = 3.4, Flavor = 3.2, Oakiness = 4.2)
predict(reduced_model, newdata = df, interval = 'predict')
anova(reduced_model)
summary(reduced_model)
```
(e) Write the final model in equation form, being careful to handle the 
qualitative predictors and interactions (if any) properly.
(f) Use the final model to predict the Quality of a wine from Region 1 with other predictors set
equal to their sample means. Also provide a 95% prediction interval for the response and a 95%
confidence interval for the mean response. Interpret the results.
```{r}
# calculating mean values
aroma_mean <- mean(wine$Aroma[wine$Region == 1], na.rm = TRUE)
print(aroma_mean)
flavor_mean <- mean(wine$Flavor[wine$Region == 1], na.rm = TRUE)
print(flavor_mean)
oakiness_mean <- mean(wine$Oakiness[wine$Region == 1], na.rm = TRUE)
print(oakiness_mean)

intercept = 6.4672
aroma_coeff = 0.5801
flavor_coeff = 1.1997
oakiness_coeff = -0.6023

predicted_val = intercept + aroma_coeff*aroma_mean + flavor_coeff*flavor_mean + oakiness_coeff*oakiness_mean
print(predicted_val)

# standard error
residualStandardError = summary(reduced_model)$sigma
SE_predict = residualStandardError * sqrt(1+sum(c((aroma_mean - mean(wine$Aroma))^2,
                                                  (flavor_mean - mean(wine$Flavor))^2,
                                                (oakiness_mean -  mean(wine$Oakiness))^2)))

# 95 percent prediction interval
df_residual <- summary(reduced_model)$df[3]
t_val = qt(0.975, df_residual)
error_margin = t_val * SE_predict
prediction_interval = c(predicted_val - error_margin, predicted_val + error_margin)
print(prediction_interval)

# 95 percent confidence interval
mean_response = SE_predict / sqrt(length(wine$Quality[wine$Region == 1]))
t_mean = qt(0.975, df_residual)
error_margin_mean = t_mean * mean_response
conf_interval_mean = c(predicted_val - error_margin_mean, predicted_val + error_margin_mean)
print(conf_interval_mean)
```
# Question 2
```{r}
diabetes <- read.csv2("C:/Users/sayem/Downloads/diabetes.csv", sep="")
str(diabetes)
summary(diabetes)
colnames(diabetes) <- gsub("\\.\\.", "", colnames(diabetes))
names(diabetes)
correlation_matrix = cor(diabetes [, c("Pregnancies", "Glucose", "BloodPressure",
"SkinThickness", "Insulin", "BMI", "DiabetesPedigreeFunction", "Age", "Outcome")])
print(correlation_matrix)
pairs(diabetes[, -10], main = "Scatterplot Matrix")


```
Part B
```{r}
lda_model = lda(Outcome ~ ., data = diabetes)
lda_predictions <- predict(lda_model, diabetes)
matrix <- table(Actual = diabetes$Outcome, Predicted = lda_predictions$class)

# Calculate sensitivity, specificity, and misclassification rate
sensitivity <- matrix[2, 2] / sum(diabetes$Outcome == 1)
specificity <- matrix[1, 1] / sum(diabetes$Outcome == 0)
misclassification_rate <- (matrix[1, 2] + matrix[2, 1]) / sum(matrix)

# Print the confusion matrix and metrics
print(matrix)
cat("Sensitivity:", sensitivity, "\n")
cat("Specificity:", specificity, "\n")
cat("Misclassification Rate:", misclassification_rate, "\n")

# Plot ROC curve
library(pROC)
rocCurve <- roc(diabetes$Outcome, lda_predictions$posterior[, 2])
par(mfrow = c(1,1))
plot(rocCurve, main = "ROC Curve for LDA", col = "blue", lwd = 2)
```
Part C
```{r}
# libraries
library(MASS)
library(pROC)

# Perform QDA
model <- qda(Outcome ~ ., data = diabetes)
# Make predictions
qda_predic <- predict(model, diabetes)
# Compute confusion matrix
conf_matrix_qda <- table(Actual = diabetes$Outcome, Predicted = qda_predictions$class)

sensitivity_qda = conf_matrix_qda[2, 2] / sum(diabetes$Outcome == 1)
specificity_qda = conf_matrix_qda[1, 1] / sum(diabetes$Outcome == 0)
misclassification_rate_qda =
(conf_matrix_qda[1, 2] + conf_matrix_qda[2, 1]) / sum(conf_matrix_qda)
# Print confusion matrix and metrics for QDA
print(conf_matrix_qda)
cat("Sensitivity (QDA):", sensitivity_qda, "\n")
cat("Specificity (QDA):", specificity_qda, "\n")
cat("Misclassification Rate (QDA):", misclassification_rate_qda, "\n")
# Plot ROC curve for QDA
roc_curve <- roc(diabetes$Outcome, qda_predict$posterior[, 2])
plot(roc_curve, main = "ROC Curve for QDA", col = "blue", lwd = 2)
```